Automatically generated NE tagged corpora for English and Hungarian

### 2012-04-27 paper/engNERwiki.tex
Changes.

Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al.~\shortcite{Nothman:08} used semi-supervised methods based on Wikipedia categories and text to classify articles into NE types. To avoid the inevitable classification errors, we obtain entity type information from the DBpedia knowledge base \cite{Bizer:09}. This enables us to have high precision information about entity types at the expense of recall: of the 3,903,467 English Wikipedia pages, 1,470,293 are covered by DBpedia.

@@

Our approach to entity identification relies on Wikipedia cross-references found in the article text. We assume that individual Wikipedia articles describe NEs. A link to an article can then be perceived as a mapping that identifies its anchor text with a particular NE.

Discovered entities are tagged with the CoNLL label assigned to them in the entity list extracted from DBpedia. If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity, and tag it as UNK. Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references. Finally, sentences with UNK links in them are removed from the corpus.

@@

\subsubsection{Nothman}

First, many pages describe common nouns (Book, Aircraft), dates or other concepts that fall outside the scope of NER. Second, not all entities are marked with links: in a Wikipedia article, typically only the first occurrence of a particular entity is linked to the corresponding page. Subsequent mentions are unmarked and often incomplete -- e.g.~family names are used instead of full names. Finally, 
...
Nothman et al.~\shortcite{Nothman:08}

@@

Nothman's solution used a trie to store the entity list. Ours uses a set, albeit with more alias types than what he used. We expect more precise tagging from our slightly more rigorous solution.

@@

To annotate our corpus with NE tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, Szeged NER corpus \cite{Szarvas:06}. It is similar to CoNLL standards: contains newswire texts, comprises ca.~200,000 tokens, and is annotated with NE class labels in line with the CoNLL annotation scheme. However, the convention of what constitutes a NE is slightly different for Hungarian.

@@

There is another special case in Hungarian: opposite to English, the number of compund words is quite large and NEs can also be subjects of compounding. In this case the common noun following the NE is joint with a hyphen, so they constitute one token. However, the joint common noun can modify the original sense of NE, depending on the semantics of the common noun. For example: in the compound \textit{Nobel-díj} ['Nobel Prize'] the common noun changes the labeling from PER to MISC, while 
%\textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but 
in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O. The solution of this problem is not obvious, and needs more investigation.

>>>>

Many authors, such as Kazama and Torisawa \shortcite{Kazama:07} and Nothman et al.~\shortcite{Nothman:08} used semi-supervised methods based on Wikipedia categories and text to classify articles into NE types. To avoid the inevitable classification errors, we obtain entity type information from the DBpedia knowledge base \cite{Bizer:09}. This enables us to have high precision information about entity types at the expense of recall: of the 3,903,467 English Wikipedia pages, 1,470,293 are covered by DBpedia\footnote{as of 18 March, 2012.}.
% TODO as of?

@@

Our approach is based on the work of Nothman et al.~\shortcite{Nothman:08}. The Wikipedia cross-references found in the article text are used to identification entities. We assume that individual Wikipedia articles describe NEs. A link to an article can then be perceived as a mapping that identifies its anchor text with a particular NE.
% TODO identifies -> ? vagy with -> as

Discovered entities are tagged with the CoNLL label assigned to them in the entity list extracted from DBpedia. If the link target is not in the entity list, or the link points to a disambiguation page, we cannot determine the type of the entity, and tag it as UNK. Links to redirect pages are resolved to point instead to the redirect target, after which they are handled as regular cross-references. Finally, sentences with UNK links in them are removed from the corpus.
%  TODO , and -> and?

@@

@@

Nothman's solution used a trie to store the entity list. Ours uses a set, albeit with more alias types than what he used. We expect more precise tagging from our slightly more rigorous solution.
% TODO: change at least use

@@

To annotate our corpus with NE tags, we chose to follow the annotation guidelines of the largest human-annotated NER corpus for Hungarian, the Szeged NER corpus \cite{Szarvas:06}. It is similar to CoNLL standards: contains newswire texts, comprises ca.~200,000 tokens, and is annotated with NE class labels in line with the CoNLL annotation scheme. However, the convention of what constitutes a NE is slightly different for Hungarian.

@@

There is another special case in Hungarian: opposite to English, the number of compund words is quite large and NEs can also be subjects of compounding. In this case the common noun following the NE is joint with a hyphen, so they constitute one token. However, the joint common noun can modify the original sense of NE, depending on the semantics of the common noun. For example: in the compound \textit{Nobel-díj} ['Nobel Prize'] the common noun changes the labeling from PER to MISC, while 
%\textit{SAP-szoftvercég} ['SAP software company'] the common noun does not change the labeling, but 
in the case of the compound \textit{WorldCom-botrány} ['WorldCom scandal'] the NE tag changes from ORG to O. The solution to this problem is not obvious, and needs more investigation.
